## ProbMod: Computational Mini Project
## Monte Carlo with Monty Hall 

# Mira


```{r,echo=FALSE}
# this function just tweaks the default behavior of the sample function. Type ?sample if you are curious to learn more
mysample <- function(x, ...) x[sample.int(length(x), ...)]
set.seed(25)
```

### Part I: Generalization of the Monty Hall Game

Consider a generalization of the famous "Monty Hall" or "Let's Make a Deal" game. There are $n\geq 3$ doors, behind one of which is a car, and behind the rest of which are goats (which you do not want). Initially, all possibilities are equally likely for the location of the car. You choose a door. Monty Hall then opens $m$ goat doors ($1\leq m \leq n-2$), and offers you the option of switching to any of the remaining $n-m-1$ doors. Assume that Monty never opens your door, and chooses with equal probabilities from all of his choices of which goat doors to open. Should you switch?


**Question 1**
   
Experiment by hand with a neighbor. Pick a value of $n$ and $m$, and play the game a few times to make sure you understand the logistics and develop your intuition about whether or not to switch.

**Question 2**
   
What is the probability of getting the car if you do not switch? How about if you switch to one of the remaining closed doors? Optional: Prove that for any valid choices of $n$ and $m$, it is always strictly better to switch.


**Question 3**

One strategy when you cannot compute a probability exactly is to simulate it by using `R` to randomly generate outcomes of the experiment, and counting the proportion of random outcomes that fall into a given event. This technique is called *Monte Carlo* simulation. It can also be used to double check an answer you computed analytically, which is what we will do in this question.

Here is a function that simulates the outcome of a classical Monty Hall game. The function takes one input variable `switch`, which should be 0 (keep your original door) or 1 (switch to one of the unopened doors).

```{r}
monty_single_game<-function(switch){
  doors<-1:3
  
  # randomly pick where the car is
  cardoor<-mysample(doors,1)
  
  # randomly pick the player's choice of door
  chosen<-mysample(doors,1)
  
  # randomly pick the door Monty opens (can't be the    player's door or the car door; when there are three   doors, there is only one door left!)
  montydoor<-mysample(doors[-c(chosen,cardoor)],1)

  # keep the door or switch?
  if (switch==1) {
    newdoor<-mysample(doors[-c(chosen,montydoor)],1)  
  }
  else{
    newdoor<-chosen
  }
  
  # determine if it is a win or loss
  if (cardoor==newdoor) return(1)
  else return(0)
}
```

Once you've executed that code chunk, you can call the function repeatedly to simulate the outcome of different games:

```{r}
switch_strategy<-0
monty_single_game(switch_strategy)
monty_single_game(switch_strategy)
monty_single_game(switch_strategy)
```

However, we can also use the function `replicate` to do this many times at once:

```{r}
num_games<-10000
switch_strategy<-0 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game(switch_strategy))
avg_win_pct<-mean(results)
print(avg_win_pct)
```

Your task is to generalize the above code (which is for $n=3$ and $m=1$) to the generalized problem considered in the previous question. First, write a function `monty_single_game_nm` that takes three input variables: `switch`, `n`, and `m`, and returns the result of a single game. Then use `replicate` to simulate 10,000 games with $n=7$ and $m=2$ in which you always switch doors. What is the approximate probability of finding the car when you switch? How do your simulation result compare to the theoretical answer you derived in the previous question?  

```{r}
monty_single_game_3<-function(switch, n=3, m=1){
  doors<-1:n
  
  # randomly pick where the car is
  cardoor<-mysample(doors,1)
  
  # randomly pick the player's choice of door
  chosen<-mysample(doors,1)
  
  # randomly pick the door Monty opens (can't be the    player's door or the car door; when there are three   doors, there is only one door left!)
  montydoor<-mysample(doors[-c(chosen,cardoor)],m)

  # keep the door or switch?
  if (switch==1) {
    newdoor<-mysample(doors[-c(chosen,montydoor)],1)  
  }
  else{
    newdoor<-chosen
  }
  
  # determine if it is a win or loss
  return (cardoor == newdoor)
}
```

```{r}
num_games<-100000
switch_strategy<-0 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_3(switch_strategy,7,2))
avg_win_pct<-mean(results)
print(avg_win_pct)

num_games<-100000
switch_strategy<-1 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_3(switch_strategy,7,2))
avg_win_pct<-mean(results)
print(avg_win_pct)
```


\
\




### Part II: Variation of the Monty Hall Game

Now, let's consider a variation where there are $n \geq 4$ doors, and the contestant gets to pick two different doors at the beginning of the game. After Monty opens $m$ ($1\leq m < n-2$) goat doors, the contestant can either keep both doors (and hope the car is behind one) or switch to just *one* of the remaining doors. 

**Question 4**

What is the probability of getting the car if you do not switch? How about if you switch to one of the remaining closed doors? Should you switch? Hint: your answer may depend on $n$ and $m$.


**Question 5**

Finally, let's write a simulation to check our answer to Question 4. Write a function `monty_single_game_nm2` that takes three input variables: `switch`, `n`, and `m`, and returns the result of a single game under these new rules. Then, for $n=7$ and $m=2$, use `replicate` to simulate 10,000 games in which you always switch doors. What is the approximate probability of finding the car when you switch? Simulate another 10,000 games in which you keep your original two doors. What is the approximate probability of finding the car when you do not switch? How do your simulation results compare to the theoretical answers you derived in the previous question?

```{r}
monty_single_game_5<-function(switch, n=3, m=1){
  doors<-1:n
  
  # randomly pick where the car is
  cardoor<-mysample(doors,1)
  
  # randomly pick the player's choice of door
  chosen<-mysample(doors,2)
  
  # randomly pick the door Monty opens (can't be the    player's door or the car door; when there are three   doors, there is only one door left!)
  montydoor<-mysample(doors[-c(chosen,cardoor)],m)

  # keep the door or switch?
  if (switch==1) {
    newdoor<-mysample(doors[-c(chosen,montydoor)],1)  
  }
  else{
    newdoor<-chosen
  }
  
  # determine if it is a win or loss
  return (cardoor %in% newdoor)
}
```

```{r}
num_games<-100000
switch_strategy<-0 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_5(switch_strategy,7,2))
avg_win_pct<-mean(results)
print(avg_win_pct)

num_games<-100000
switch_strategy<-1 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_5(switch_strategy,7,2))
avg_win_pct<-mean(results)
print(avg_win_pct)
```


```{r}
num_games<-100000
switch_strategy<-0 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_5(switch_strategy,10,3))
avg_win_pct<-mean(results)
print(avg_win_pct)

num_games<-100000
switch_strategy<-1 # choose 0 or 1 here
results<-replicate(num_games,monty_single_game_5(switch_strategy,10,3))
avg_win_pct<-mean(results)
print(avg_win_pct)
```







